---
title: "USA Computer and Internet Accessibility in 2022"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(dplyr)
library(readr)
library(tidyverse)
library(kableExtra)
library(tidyr)
library(sf) # Map
```

```{r}
# Prepared data for graphs from Education_Comp_Use file

Edu_Comp_22 <- read_csv("Data/Edu_Comp_22.csv", show_col_types = FALSE)
Code_book <- read_csv("Data/Code Book.csv", show_col_types = FALSE)

data <- read.csv("Data/Edu_Comp_22.csv", header = FALSE, skip = 2) # Skip first two rows

col_names <- read.csv("Data/Edu_Comp_22.csv", header = FALSE, nrows = 1) # Read only the first row as column names
colnames(data) <- col_names # Assigns row 1 header to data

cols_to_remove <- grep("M$", col_names) 
data <- data[, -cols_to_remove, drop = FALSE]

new_col_names <- Code_book[[3]] # Extract the third column

colnames(data) <- new_col_names

data <- data[, -ncol(data), drop = FALSE]

data <- data %>% # Putting states in economic regions
  mutate(
    Region = case_when(
      State %in% c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont") ~ "New England",
      State %in% c("Delaware", "District of Columbia", "Maryland", "New Jersey", "New York", "Pennsylvania") ~ "Mideast",
      State %in% c("Illinois", "Indiana", "Michigan", "Ohio", "Wisconsin") ~ "Great Lakes",
      State %in% c("Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota") ~ "Plains",
      State %in% c("Alabama", "Arkansas", "Florida", "Georgia", "Kentucky", "Louisiana", "Mississippi", "North Carolina", "South Carolina", "Tennessee", "Virginia", "West Virginia") ~ "Southeast",
      State %in% c("Arizona", "New Mexico", "Oklahoma", "Texas") ~ "Southwest",
      State %in% c("Colorado", "Idaho", "Montana", "Utah", "Wyoming") ~ "Rocky Mountain",
      State %in% c("Puerto Rico") ~ "Puerto Rico",
      State %in% c("Alaska", "California", "Hawaii", "Nevada", "Oregon", "Washington") ~ "Far West",
      TRUE ~ NA_character_
    )
  )

data <- data %>% # Calculating the proportion those without a computer and internet subscription for each level of education
  mutate(
    Prop_no_Comp = (`< HS no Comp`+ `< HS w/Comp w/out IS`+ `HS no Comp`+ `HS some Col w/Comp w/out IS`+ `BA+  no Comp` + `BA+  w/Comp w/out IS`)/`State Total`
  )

# New columns for proportion of people with no internet access or no computer by education level, that is a conditional probability, P(no_Comp|education level)
data <- data %>%
  mutate(
    lths_Prop_no_Comp = ((`< HS no Comp`+ `< HS w/Comp w/out IS`)/`< HS Total`), 
    hs_Prop_no_Comp = ((`HS no Comp`+ `HS some Col w/Comp w/out IS`)/`HS some Col Total`),
    ba_Prop_no_Comp = ((`BA+  no Comp` + `BA+  w/Comp w/out IS`)/`BA+ Total`))

summary_data <- data %>%
  group_by(Region) %>%
  summarise(
    mean_no_Comp = mean(Prop_no_Comp),
    median_no_Comp = median(Prop_no_Comp)
  )

# Prepare data for Boxplot No Computer or Internet Subscription by Education level
edu_level_data <- data %>%
  select(Prop_no_Comp, lths_Prop_no_Comp, hs_Prop_no_Comp, ba_Prop_no_Comp
  )
# Rename columns to more understandable
colnames(edu_level_data) <- c("USA", "Less Than High School", "HS and Some College", "Bachelors Plus")

# Table
Summary_Edu_level <- summary(edu_level_data) %>%
  kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# Long format for plotting
edu_level_data_long <- tidyr::gather(edu_level_data, key = "Variable", value = "Proportion") # Wide to long

# Reorder the levels of education for x axis
edu_level_data_long$Variable <- factor(edu_level_data_long$Variable, 
                                       levels = c("Less Than High School", "HS and Some College", "Bachelors Plus", "USA"))
```
```{r}
# Data frame to send to Professor so he can show us how to merge and use spatial operation with the USA shape file
data_for_map <- data.frame(State = data$State,
                       Region = data$Region,
                       Prop_no_Comp = data$Prop_no_Comp)

# Specify the file path where you want to save the CSV file
file_path <- "Data/data_for_map.csv"

# Write the data frame to a CSV file
#<<<<<<< HEAD
#write.csv(data_for_map, file_path, append = TRUE, row.names = FALSE)
#=======
#write.csv(data_for_map, file_path, row.names = FALSE)
#>>>>>>> 1fae9b3f48ecb0a4c9b55b042f32a933267a25d1

```

```{r}
# 1 Categorical Plot Ascending Order
ascending_data <- summary_data[order(summary_data$mean_no_Comp), ]
ascending_data$Region <- factor(ascending_data$Region, levels = ascending_data$Region[order(ascending_data$mean_no_Comp)])

# Plot with reordered levels
Econ <- ggplot(ascending_data, aes(x = Region, y = mean_no_Comp)) +
  geom_bar(stat = "identity", fill = "darkblue") +
  labs(x = "", y = "") +
  labs(title = "No Computer or Internet Subscription", x = "", y = "") + 
  labs(subtitle = "By Economic Region") +
  labs(caption = "Source: Census.gov/data") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, .18, by = 0.05))
  
```

```{r}
# 2 Numerical Plot
box_Edu_Level <- ggplot(edu_level_data_long, aes(x = Variable, y = Proportion, fill = Variable)) +
  geom_boxplot(position = position_dodge(width = 0.75), width = 0.5, alpha = 0.5) +
  labs(title = 'No Computer or Internet Subscription',
       x = '',
       y = '',
       fill = 'Education Level') +
  labs(subtitle = "By Education Level") +
  labs(caption = "Source: Census.gov/data") + 
  scale_fill_manual(values = c("#FF0000", "white", "darkblue", "skyblue")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = FALSE) + # Remove legend
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, .4, by = 0.1))
```

```{r}
# 3 For Bivariate Cat-Cat Plot, Education Level and No Computer in Internet Subscription
# Rename columns
data_renamed <- data %>%
  rename(
    "Less Than High School" = lths_Prop_no_Comp,
    "High School and Some College" = hs_Prop_no_Comp,
    "Bachelor's Plus" = ba_Prop_no_Comp
  )
# Pivot longer
data_long_edu_level <- pivot_longer(data_renamed, 
                                    cols = c("Less Than High School", "High School and Some College", "Bachelor's Plus"), 
                                    names_to = "Education_Level", 
                                    values_to = "Percent")

# Reorder horizontal axis to match Econ graph, levels of "Region" in data_long_edu_level to match ascending_data

data_long_edu_level$Region <- factor(data_long_edu_level$Region, levels = ascending_data$Region)

# Graph
c_c <- ggplot(data_long_edu_level, aes(x = Region, y = Percent, fill = Education_Level)) + 
  geom_bar(position = "fill", stat = "identity") +
  labs(title = "No Computer or Internet Subscription", x = "", y = "") + 
  labs(subtitle = "By Economic Region & Education Level") +
  labs(caption = "Source: Census.gov/data") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, by = 0.2)) +
  scale_fill_manual(values = c("#FF0000", "#FFFFFF", "darkblue"))

```

```{r}
# Map

# rm(list = ls()) # Clear workspace
usaMap=read_sf("Data/gz_2010_us_040_00_500k.json")
# ggplot(data = usaMap)+ geom_sf() +coord_sf(crs = st_crs(2163)) 

variables=read.csv("Data/data_for_map.csv")
# sort(usaMap$NAME)==sort(variables$State)

# Merge
usaMap=merge(usaMap,variables, by.x='NAME', by.y='State')
# usaMap
```

```{r}
# Map continued

usaMap %>%
  group_by(Region) %>%
  summarise_at(vars(Prop_no_Comp), list(Prop_no_Comp_mean=mean)) -> variablesByRegion

# New Map
# variablesByRegion

Map_Econ <- ggplot(variablesByRegion) + 
  geom_sf(aes(fill=Prop_no_Comp_mean)) + 
  coord_sf(crs = st_crs(2163)) +
  labs(title = "No Computer or Internet Subscription") +
  labs(subtitle = "Source: Census.gov/data") + 
  scale_fill_gradient(name = "Percent", low = "skyblue", high = "darkblue")
```
## Sidebar {.sidebar}

##Column {.tabset .tabset-fade data-width=350}
-----------------------------------------------------------------------

### Which regions have the highest percent without computers or internet subscription? 

```{r}
Econ
```


### Which region has the highest discrepency of technology accessibility by education level? 

```{r}
c_c 
```

##Column {data-width=350}
#-----------------------------------------------------------------------

### What level of education has the most accessibility to technology?

```{r}
box_Edu_Level

```

### Of the lower 48 states, which region has the least accessiblity to technology? 

```{r}
Map_Econ

```

