---
title: "690V Data Visualization Practice"
author: "Group 2"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
---
```{r}
# install.packages("reader")
# install.packages("reshape2")

```

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Import Data Hello

```{r setup, include=FALSE}
Edu_Comp_22 <- read_csv("Data/Edu_Comp_22.csv")
Code_book <- read_csv("Data/Code Book.csv")
```

## Prepare Data

```{r}
# Read the data set 
data <- read.csv("Data/Edu_Comp_22.csv", header = FALSE, skip = 2) # Skip first two rows
col_names <- read.csv("Data/Edu_Comp_22.csv", header = FALSE, nrows = 1) # Read only the first row as column names
colnames(data) <- col_names # Assigns row 1 header to data
```

```{r}
# Remove columns with margin of error
cols_to_remove <- grep("M$", col_names) 
data <- data[, -cols_to_remove, drop = FALSE]

# Remove last NA column
data <- data[, -ncol(data), drop = FALSE]
```

# Replace the headers with new_col_names
colnames(data) <- new_col_names
```

## Categorical Plot 

```{r}
# Categorial states in regions 
data <- data %>%
  mutate(
    Region = case_when(
      State %in% c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont") ~ "New England",
      State %in% c("Delaware", "District of Columbia", "Maryland", "New Jersey", "New York", "Pennsylvania") ~ "Mideast",
      State %in% c("Illinois", "Indiana", "Michigan", "Ohio", "Wisconsin") ~ "Great Lakes",
      State %in% c("Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota") ~ "Plains",
      State %in% c("Alabama", "Arkansas", "Florida", "Georgia", "Kentucky", "Louisiana", "Mississippi", "North Carolina", "South Carolina", "Tennessee", "Virginia", "West Virginia") ~ "Southeast",
      State %in% c("Arizona", "New Mexico", "Oklahoma", "Texas") ~ "Southwest",
      State %in% c("Colorado", "Idaho", "Montana", "Utah", "Wyoming") ~ "Rocky Mountain",
      State %in% c("Puerto Rico") ~ "Puerto Rico",
      State %in% c("Alaska", "California", "Hawaii", "Nevada", "Oregon", "Washington") ~ "Far West",
      TRUE ~ NA_character_
    )
  )
```

```{r}
# New column for proportion of people with no computer or internet access
data <- data %>%
  mutate(
    Prop_no_Comp = (`< HS no Comp`+ `< HS w/Comp w/out IS`+ `HS no Comp`+ `HS some Col w/Comp w/out IS`+ `BA+  no Comp` + `BA+  w/Comp w/out IS`)/`State Total`
  )
```

```{r}
# New columns for proportion of people with no internet access or no computer by education level, that is a conditional probability, P(no_Comp|education level)
data <- data %>%
  mutate(
    lths_Prop_no_Comp = ((`< HS no Comp`+ `< HS w/Comp w/out IS`)/`< HS Total`), 
    hs_Prop_no_Comp = ((`HS no Comp`+ `HS some Col w/Comp w/out IS`)/`HS some Col Total`),
    ba_Prop_no_Comp = ((`BA+  no Comp` + `BA+  w/Comp w/out IS`)/`BA+ Total`))
```
## Bar Graph

```{r}
ggplot(data, aes(Region)) +
  geom_bar(fill ="skyblue") +
  labs(title = "Bureau of Economic Analysis Regions",
       y = "Number of States") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, 15, by = 1))


```
```{r}
summary_data <- data %>%
  group_by(Region) %>%
  summarise(
    mean_no_Comp = mean(Prop_no_Comp),
    median_no_Comp = median(Prop_no_Comp)
  )

# Create the bar plot
ggplot(summary_data, aes(x = Region, y = mean_no_Comp)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "Bureau of Economic Analysis Regions", y = "Proportion") +
  ggtitle("USA: No Computer or Internet Access") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
## Sample Histogram

```{r}
ggplot(data, aes(x = `State Total`)) +
  geom_histogram(binwidth = 1000000, fill = "skyblue", color = "black") +
  labs(title = "State Population Distribution",
       x = "Population in Millions",
       y = "Number of States") +
  scale_y_continuous(breaks = seq(0, 14, by = 2),
                     labels = function(x) format(x, scientific = FALSE)) +
  scale_x_continuous(breaks = seq(0, max(data$`State Total`), by = 5e6),
                     labels = function(x) format(x / 1e6, scientific = FALSE, big.mark = ",")) +
  theme_minimal()

```


```{r}
ggplot(data, aes(x = Prop_no_Comp)) +
  geom_histogram(bins = 50, fill = 'red', color = 'black') +
  labs(title = 'No Computer or Internet Access Distribution by Region',
       x = 'Proportion',
       y = 'Number of States') +
  facet_wrap(~ Region)

```


```{r}
ggplot(data, aes(y = Prop_no_Comp)) +
  geom_boxplot(fill='skyblue') +
  labs(title = 'USA:No Computer or Internet Access',
       y = 'Proportion') +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```
```{r}
# Boxplot No Computer or Internet Access by Education level
edu_level_data <- data %>%
  select(Prop_no_Comp, lths_Prop_no_Comp, hs_Prop_no_Comp, ba_Prop_no_Comp
  )
# Rename columns to more understandable
colnames(edu_level_data) <- c("USA", "Less Than High School", "HS and Some College", "Bachelors Plus")

summary(edu_level_data)
```
```{r}
# Long format for plotting
edu_level_data_long <- tidyr::gather(edu_level_data, key = "Variable", value = "Proportion") #convert wide to long

# Reorder the levels of education for x axis
edu_level_data_long$Variable <- factor(edu_level_data_long$Variable, 
                                       levels = c("Less Than High School", "HS and Some College", "Bachelors Plus", "USA"))

# Create boxplot
ggplot(edu_level_data_long, aes(x = Variable, y = Proportion, fill = Variable)) +
  geom_boxplot(position = position_dodge(width = 0.75), width = 0.5, alpha = 0.5) +
  labs(title = 'USA: No Computer or Internet Access by Education Level',
       x = '',
       y = 'Proportion',
       fill = 'Education Level') +
  scale_fill_manual(values = c("skyblue", "navy", "white", "red")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = FALSE) # Remove legend

```



